### 2024-02-19 chap_case_study/chap_case_study.tex
Merge taxonomy paper + Implement Hervé's comments on case_study

As detailed by \citet{limpens2024pathway}, the analysis carried out in this work can be applied to any regional whole-energy system. As a densely-populated and highly-industrialised country with limited local renewable potentials (\ie mainly solar and wind representing up to 50\% of the primary mix by 2050), the transition of Belgium from a fossil-dominated system in 2020 (Appendix \ref{app:bel_2020}) to carbon-neutrality in 2050 makes it an intricate case study. Moreover, this case study and the subsequent analyses can be transferred - to some extent - to other industrialised countries highly dependent on fossil fuels with limited local renewable potentials (\eg the Netherlands or Germany) \cite{dommisse2020modelling}. This chapter presents the different demands to satisfy, with a particular focus on the non-energy demand, as well as the resources available and the conversion technologies to supply those. For a comprehensive understanding and detailed descriptions of the technologies, please refer to the documentation \cite{readthedocs_pathway}. Then, the uncertainty ranges considered for some of the parameters are detailed. Finally, the \ce{CO2}-budget over the 2020-2050 transition is presented.

\section*{Contributions}
\label{sec:cs:contributions}
First, as pointed out by \citet{rixhon2022integration}, where most of the studies assessing whole-energy system integrate energy demands (\ie electricity, heart and mobility), the \acrfull{NED} is often not considered. The latter is defined as ‘\textit{energy products used as raw materials in the different sectors; that is not consumed as a fuel or transformed into another fuel}’ \cite{Eurostat2019}. The previous analyses carried out with \gls{ESTD} considered the non-energy demand as the related primary energy needs, \ie either natural gas or \gls{LFO}. To minimise the total cost of the system, the model simply selected the cheapest between the two resources (\ie natural gas). This work goes one step further and accounts for the \gls{NED} as a demand of three commodities (\ie \gls{HVC}, ammonia and methanol) as well as with associated production technologies. This allows bringing the non-energy sector to a similar level of details as the other sectors.  Keeping the same methodology to define \gls{EUD} in EnergyScope as \citet{Limpens2020}, this work considers updated values given the latest release of the \og EU reference scenario 2020 : energy, transport and GHG emissions: Trends to 2050 \fg by the European Commission \cite{EuropeanCommission2021}.

Second, given the focus of this work on the electrofuels, the case study includes a more explicit representation of them: where the previous definition of the case study considered only renewable hydrogen and methane \cite{limpens2021generating}, now, e-ammonia and e-methanol (as well as their fossil-based equivalents), are implemented in the case study. As detailed later on, these electrofuels are considered as renewable in the sense that their \gls{GWP} is zero. This more explicit representation of the molecules themselves also comes with a more exhaustive integration of the ways to produce and use them in the system. For instance, considering the case of ammonia, on the top of the import routes, the Haber-Bosch process ,on one hand, is accounted for to supply this molecule. On the other hand, ammonia-driven \gls{CCGT} or ammonia-cracking-to-hydrogen are included as ways to consume it.

Third, as nuclear energy could be a real game-changer in the energy transition worldwide \cite{IEA2022nuclear}, and especially in Belgium, this thesis has integrated the uncertain potential to install \gls{SMR} from 2040 onward.

Fourth, where previous works considered a prescribed \ce{CO2} trajectory to reach carbon-neutrality by 2050 \cite{limpens2021generating, limpens2024pathway}, the case study analysed in this thesis is subject to a \ce{CO2}-budget for the transition, \ie limiting the total amount of emissions over the transition. Based on the estimated world budget provided by the \gls{IPCC} to limit the global warming to +1.5°C by the end of the century, the grandfathering approach has been used to allocate part of this budget to the Belgian energy transition.

Finally, to a smaller extent, this work includes updated values for some parameters compared to the work of \citet{limpens2021generating}. The main change concerns the cost and performance of private mobility vehicles, which is specifically a key components in the European \cite{biresselioglu2018electric} and Belgian \cite{BFP_mob} energy transitions. Based on the work of \citet{national2013transitions}, \citet{limpens2021generating} excessively favoured \gls{FC} car versus \gls{BEV}. Besides their lower efficiencies (\ie about 50\% less efficient), \gls{FC} car had the lion share of the private mobility, compared to \gls{BEV} given the higher CAPEX (\ie up to 10\% more expensive) and limited range (\ie 24kWh battery) of the latter, in previous works \gls{limpens2021generating,rixhon2021role}. In these, the more limited potential to import electricity from abroad and produce it locally via \gls{VRES} forced the model to rather electrify the low-temperature heat sector rather than the private mobility running on supposedly infinitely available renewable hydrogen. To align with other similar works on the modelling of whole-energy system \cite{schnidrig2021modelling, EuropeanCommission2021}, the CAPEX and efficiency of fuel cell cars have been increased. Regarding \gls{BEV}, while the CAPEX has been kept unchanged, the efficiency and the battery capacity, \ie the range, have been increased. As seen in the results, this change of data made \gls{BEV} often more competitive than its hydrogen-based equivalent.

\section{End-use demands}
\label{sec:cs:demand}
End-use demands, exogenously imposed as inputs to the model, are characterised by yearly quantities to satisfy and are also distributed over the different hours of each representative years of the transition, in order to account for their daily or seasonal variability \cite{Limpens2020,limpens2021generating}. In this work, the yearly end-use demands (EUD) for all sectors are calculated from the forecast proposed by the European Commission for Belgium (Appendix 2 in report \cite{EuropeanCommission2021}).

\subsection{Non-energy demand}
\label{subsec:cs:NED}
The \gls{NED} currently represents around 20\% of the final energy consumption in Belgium \cite{FPSEconomy2021}.  This section summarises the rationale of adding a higher level of details to the\gls{NED} compared to what was done in the previous version of the case study \cite{limpens2021generating}. Then, it explains the methodology used to quantify this demand.\\

@@

Over the recent history, there has been a relatively constant evolution of three main categories of the final consumption for non-energy use in Belgium, \cite{statbel_NED_2019}: (i) naphtha and \gls{LPG} (between 59\% and 67\% of the total final consumption, around 59.4\,TWh in 2019), (ii) fossil gas (between 9\% and 14\%, 11.8\,TWh in 2019), and (iii) others (\ie bitumen, coal tar and other oil products) (between 21\% and 28\%). Naphtha and \gls{LPG} are consumed in a naphtha cracker, which results in ethylene and propylene, what will be considered as \gls{HVC} in the rest of this work. Similarly, fossil gas, as non-energy carrier, is used in steam methane reformer to produce the required hydrogen to the synthesis process of ammonia. The small shares of bitumen and coal tar are respectively used for roadworks and to produce synthetic gas through gasification. Finally \og other oil products'' take into account, indistinguishably, tar and sulphur as well as by-products of the refineries (\eg \gls{BTX}). About methanol, there is currently no production plant in Belgium even if the country plays a role in trading this commodity between its neighbouring countries and consumes part of what it imports.\\

@@

\noindent
The non-energy demand studied in this analysis focuses on the chemical industry (more than 90\% of the non-energy use in Belgium) and, similarly to other studies \cite{IEA2018_petrochemicals, daioglou2014energy}, is split between the three aforementioned main groups of products (\ie \gls{HVC}, ammonia and methanol). Before describing these three demands, this study excludes bitumen, coal tar and \og other oil products\fg. The first two represent marginal shares of the current non-energy use (\ie 4\% and 5\% respectively) in such a way that they should not affect the conclusions provided by this study. As described previously, the latest are mostly by-products from refineries that the system uses because they are available. However, in a perspective of defossilisation, since the future of fossil-based refineries is unclear, they have not been implemented in this study nor their by-products.

Regarding \gls{HVC}, the future of their production is highly uncertain. Because of the new regulations and strategies promoting recycling and limitation of single-use plastics \cite{EU_plastics}.  Besides this uncertainty, Belgium stays a major exporter as approximately 2/3 of plastic raw materials produced locally are exported abroad \cite{agoria_plastics}. Even if a significant part of \gls{HVC} produced in Belgium is not locally consumed, this demand has been set based on the assumption that Belgium will keep its industrial activity in this sector. In other words, this assumption does not deduce the part of local production  of \gls{HVC} being then exported (and not consumed locally), unlike ammonia and methanol, which will be more traded commodities in the future (as energy carriers and non-energy products). Therefore, the actual demand of \gls{HVC} is inferred from the consumption of naphtha and \gls{LPG} as non-energy use as well as energy-carrier in the chemical and petrochemical industries \cite{statbel_NED_2019}. This assumption is based on the fact that, in the conversion processes to produce \gls{HVC} from naphtha or \gls{LPG}, these fuels also serve as energy-carrier to supply the process itself. Then, given the respective efficiencies (1.83t$_{\text{naphtha}}$/t$_{\text{HVC}}$ and 1.67t$_{\text{LPG}}$/t$_{\text{HVC}}$) \cite{IEA2018_petrochemicals}, the current demand of \gls{HVC} is estimated equal to 3069\,kt, without making distinctions between the different chemicals (\ie ethylene, propylene and \gls{BTX}).

The ammonia sector in Belgium is quite different: the country locally produces and imports ammonia much more than it exports it. Thanks to a database from the United Nations \cite{UN_statistics} and the National Bank of Belgium, it has been identified that, over the last ten years, Belgium has imported 1010\,kt of ammonia, exported 105\,kt and locally produced 990\,kt on average. Therefore, on top of the local production, the net import (\ie import minus export) is also included in this non-energy demand. This gives a current demand of 1895\,kt of ammonia.

Concerning the demand of methanol, similarly to ammonia, this work solely considers the net imports as there is no local production in Belgium. To define the actual non-energy demand of methanol, only a 51\%-share of this net import is kept since, according to the Methanol Institute and \gls{MMSA}, this share is used for formaldehyde production in Belgium \cite{MMSA51}. Currently, the rest of the methanol is used for energy purposes, mostly as \gls{MTBE} in gasoline blending. This methodology gives a current non-energy demand of methanol of 269\,kt.

Eventually, after converting these masses of products into energy content (\ie LHV: \gls{HVC} - 47\,MJ/kg, ammonia - 18.8\,MJ/kg and methanol - 19.9\,MJ/kg), two final assumptions are made: (i) the current shares of each of the three commodities (\ie 77.9\%, 19.2\% and 2.9\% for \gls{HVC}, ammonia and methanol, respectively) are supposed to remain unchanged over the transition; (ii) the overall \gls{NED}, in absolute terms, follows the growing rate presented by \citet{EuropeanCommission2016}, \ie around 7\% between 2020 and 2050.

\subsection{Forecast of the demands over the transition}
\label{subsec:cs:EUD_forecast}
In their latest report \cite{EuropeanCommission2021}, the European Commission forecasts a significant and abrupt increase of the \gls{NED} compared to their previous report \cite{EuropeanCommission2016}, \ie +80\% over the 2020-2030 time window. Given this discrepancy that is unsubstantiated and specific to the case of Belgium, the evolution trend of the \gls{NED} of the current work has been inferred from the previous edition, published in 2016, \cite{EuropeanCommission2016}. Between 2020 and 2050, one observes a noteworthy increase of the electricity (+40\%), passenger (+45\%) and freight mobility (+35\%) demands (see \Cref{fig:cs_demands}). The rise of the non-energy demand is more limited, \ie +6\%, whereas the heating demands is forecast to decrease: -11\% for the low-temperature heat demand and -3\% for the high-temperature heat demand. This is explained by a better insulation of buildings and an improved efficiency of industrial processes. Regarding the center graph of \Cref{fig:cs_demands}, it is the aggregation of the same data as in the left graph but per category, rather than per sector, with the non-energy demand being associated with the industry. This illustrates how industrialised is Belgium, compared to households and services, and, consequently, highly energy-intensive. The right graph of \Cref{fig:cs_demands} gives the passenger and the freight mobility. The sharp increase from 2020 to 2025 is due the COVID-crisis that has significantly reduced these demands in 2020. As far as the hourly discretisation of these demands is concerned, time series are based on historical values of 2015 for the fluctuating parts of electricity and low-temperature heating demands \cite{Limpens2020}. A daily time series is used for the passenger mobility and applied similarly to every days of the year. Finally, for the other demands (\ie high-temperature heat, freight mobility, \gls{NED} and constant share of electricity and low-temperature heat demands), the yearly demand is distributed uniformly over the different hours of the year.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.49\textwidth]{EUD_sec.pdf}
\includegraphics[width=0.49\textwidth]{EUD_cat.pdf}
\includegraphics[width=0.49\textwidth]{EUD_mob.pdf}
\caption{EnergyScope splits the whole-energy system end-use demands (EUD) into two sets: (non)-energy and transport-related. This figure presents the nominal values of each of these demands. In the center graph, the non-energy demand has been fully associated with the industrial demand. As detailed previously, the non-energy demand is expressed in tons of physical products (\ie \glsxtrfull{HVC}, ammonia and methanol) and then translated into their respective energy equivalent, in TWh.}
\label{fig:cs_demands}
\end{figure}

\section{Resources}
\label{sec:cs:resources}
To supply the aforementioned demands, EnergyScope Pathway implements a variety of resources defined by their cost of purchasing, $\mathit{c}_{\mathrm{op}}$, their global warming potential, $\emph{gwp}_{\mathrm{op}}$, as well as their 
availability, as detailed by \citet{limpens2024pathway}. First, the evolution of the respective costs of purchasing is presented (see \Cref{fig:cs_resources_cost}). Regarding ``renewable electrofuels'', these are in line with the recent study of \citet{genge2023supply} who carried out an extensive review and ``meta-analysis\cite{grant2009typology,page2021prisma} of 30 studies on the supply costs of chemical energy carriers''. Then, besides their cost, the resources are either limited or unlimited in terms of availability and either renewable or not. The limitation in terms of availability can be direct or indirect. On the one hand, woody (23.4\,TWh) and wet biomass (38.9\,TWh) are arbitrarily limited by their local potentials and the consumption of waste (17.8\,TWh) and coal (33.4\,TWh) is assumed not to exceed the current use. On the other hand, wind, solar, hydro and uranium are limited by the technical potentials of \gls{PV} panels (59.2\,GW), onshore (10\,GW) and offshore (6\,GW) wind turbines, run-of-the-river power plants (0.1\,GW) and nuclear power plants (6\,GW), respectively. In line with the work of \citet{PATHS2050} and the maximum capacity of conventional nuclear reactors that have been installed in Belgium, the same 6\,GW are assumed to be the maximum capacity for \gls{SMR}. Imported electricity is limited in two ways: the potential of instantaneous capacity of interconnection with neighbouring countries (\ie 11.9\,GW by 2050 \cite{ELIA_2050}) and a limitation to 30\% of the yearly electricity end-use demand (\ie 32.4\,TWh by 2050) \cite{limpens2021generating}. In the current work, the electrofuels (\ie e-methane, e-hydrogen, e-methanol and e-ammonia) are assumed to be ``sustainable" in the sense that they do not increase the concentration of \ce{CO2} in the atmosphere \cite{rixhon2021terminology}. In practice, it means that their \gls{GWP} is assumed to be zero in the model. Regarding specifically these electrofuels, the \citet{h2coalition} has carried out an extensive techno-economic analysis to estimate their respective cost of purchasing, after having identified some key locations from which importing these energy carriers (\eg Chile, Australia or Morocco). As the amount to import from each of these locations is hard to forecast, the current work considers the average cost between the different locations. Besides these, every other resource has its specific \gls{GWP} like coal ($\emph{gwp}_{\mathrm{op,coal}}=0.40$\,kt$_{\ce{CO2},\text{eq}}$/GWh), natural gas ($\emph{gwp}_{\mathrm{op,NG}}=0.27$\,kt$_{\ce{CO2},\text{eq}}$/GWh) or the fossil-based molecules equivalent to the electrofuels (\eg $\emph{gwp}_{\mathrm{op,ammonia}}=0.46$\,kt$_{\ce{CO2},\text{eq}}$/GWh or $\emph{gwp}_{\mathrm{op,methanol}}=0.41$\,kt$_{\ce{CO2},\text{eq}}$/GWh).

@@

\includegraphics[width=0.49\textwidth]{Res_ren.pdf}
\includegraphics[width=0.49\textwidth]{Res_foss.pdf}
\includegraphics[width=0.49\textwidth]{Res_others.pdf}
\caption{Cost of purchasing the different resources. Besides the free local renewables (\ie sun, wind and hydro) limited by technical potentials, EnergyScope accounts for renewable energy carriers and their respective fossil counterparts (left and center graphs). These fuels can be imported from abroad without limitation on their availability. Other carriers are limited either by their local potentials (\ie biomass and waste) or other considerations like the power grid interconnections or the capacity of nuclear power plants.}
\label{fig:cs_resources_cost}
\end{figure}

\section{Conversion technologies}
\label{sec:cs:technologies}
As the end-use demands are defined as energy (and non-energy with the \gls{NED}) services rather than a certain quantity of oil or solar irradiance, for instance, technologies are implemented to convert these resources into the end-use demands. Besides their CAPEX, OPEX and lifetime defined in Section \ref{sec:meth:ES}, production and conversion technologies (\ie \gls{CCGT}, car or boiler) have a conversion efficiency whereas storage technologies (\ie thermal storage, battery or molecule storage) exhibit their own charge/discharge losses. There are also infrastructure technologies. They encompass, for instance, the power grid, the \gls{DHN} or technologies to produce intermediate energy carriers (\eg wood pyrolysis, biomethanolation or steam methane reforming to produce hydrogen). Not digging into too much details about the exhaustive list of these technologies presented in previous works \cite{limpens2021generating}, this section rather focuses on the implementation of \glsxtrfull{SMR} then the technologies to supply the \glsxtrfull{NED}.

@@

For further details on these analyses, the interested reader is invited to refer to aforementioned for further published works.

@@

The exhaustive list of the parameters accounted in this work is presented in Appendix \ref{app:UC_full}.

>>>>

As detailed by \citet{limpens2024pathway}, the analysis carried out in this work can be applied to any regional whole-energy system. As a densely-populated and highly-industrialised country with limited local renewable potentials (\ie mainly solar and wind representing up to 50\% of the primary mix by 2050), the transition of Belgium from a fossil-dominated system in 2020 (Appendix \ref{app:bel_2020}) to carbon-neutrality in 2050 makes it an intricate case study. Moreover, this case study and the subsequent analyses can be transferred - to some extent - to other industrialised countries highly dependent on fossil fuels with limited local renewable potentials (\eg the Netherlands or Germany) \cite{dommisse2020modelling}. This chapter presents the different demands to satisfy, with a particular focus on the non-energy demand, as well as the resources available and the conversion technologies to supply these demands. For a comprehensive understanding and detailed descriptions of the technologies, please refer to the documentation \cite{readthedocs_pathway}. Then, the uncertainty ranges considered for some of the parameters are detailed. Finally, the \ce{CO2}-budget over the 2020-2050 transition is presented.

\section*{Contributions}
\label{sec:cs:contributions}
First, as pointed out by \citet{rixhon2022integration}, where most of the studies assessing whole-energy system integrate energy demands (\ie electricity, heat and mobility), the \acrfull{NED} is often not considered. The latter is defined as ‘\textit{energy products used as raw materials in the different sectors; that is not consumed as a fuel or transformed into another fuel}’ \cite{Eurostat2019}. The previous analyses carried out with \gls{ESTD} considered the non-energy demand as the related primary energy needs, \ie either natural gas or \gls{LFO}. To minimise the total cost of the system, the model simply selected the cheapest between the two resources (\ie natural gas). This work goes one step further and accounts for the \gls{NED} as a demand for three commodities (\ie \gls{HVC}, ammonia and methanol) as well as the associated production technologies. This allows bringing the non-energy sector to a similar level of details as the other sectors.  Keeping the same methodology to define \gls{EUD} in EnergyScope as \citet{Limpens2020}, this work considers updated values given the latest release of the \og EU reference scenario 2020 : energy, transport and GHG emissions: Trends to 2050 \fg by the European Commission \cite{EuropeanCommission2021}.

Second, given the focus of this work on the electrofuels, the case study includes a more detailed representation of them: where the previous definition of the case study considered only renewable hydrogen and methane \cite{limpens2021generating}, now, e-ammonia and e-methanol (as well as their fossil-based equivalents), are implemented in the case study. As detailed later on, these electrofuels are considered as renewable in the sense that their \gls{GWP} is zero. This more detailed representation of the molecules themselves also comes with a more exhaustive integration of the ways to produce and use them in the system. For instance, considering the case of ammonia, on the top of the import routes, on the one hand, the Haber-Bosch process is accounted for to supply this molecule. On the other hand, ammonia-driven \gls{CCGT} or ammonia-cracking-to-hydrogen are included as ways to consume it.

Third, where previous works considered a prescribed \ce{CO2} trajectory to reach carbon-neutrality by 2050 \cite{limpens2021generating, limpens2024pathway}, the case study analysed in this thesis is subject to a \ce{CO2}-budget for the transition, \ie limiting the total amount of emissions over the transition. Based on the estimated world budget provided by the \gls{IPCC} to limit the global warming to +1.5°C by the end of the century, the grandfathering approach has been used to allocate part of this budget to the Belgian energy transition.

Fourth, as nuclear energy could be a real game-changer in the energy transition worldwide \cite{IEA2022nuclear}, and especially in Belgium, this thesis has integrated the uncertain decision to install \gls{SMR} from 2040 onward.

Finally, this work includes updated values for some parameters compared to the work of \citet{limpens2021generating}. The main change concerns the cost and performance of private mobility vehicles, which is a key components in the European \cite{biresselioglu2018electric} and Belgian \cite{BFP_mob} energy transitions. Based on the work of the \citet{national2013transitions}, \citet{limpens2021generating} excessively favoured \gls{FC} car versus \gls{BEV}. In previous works \cite{limpens2021generating,rixhon2021role}, despite their lower efficiencies (\ie about 50\% less efficient), \gls{FC} cars had the lion share of the private mobility compared to battery electric vehicles(BEV) given the higher CAPEX (\ie up to 10\% more expensive) and limited range (\ie 24kWh battery) of the latter. In these, the more limited potential to import electricity from abroad and produce it locally via \gls{VRES} forced the model to rather electrify the low-temperature heat sector rather than the private mobility running on supposedly infinitely available renewable hydrogen. To align with other similar works on the modelling of whole-energy systems \cite{schnidrig2021modelling, EuropeanCommission2021}, the CAPEX and efficiency of fuel cell cars have been increased. Regarding \gls{BEV}, while the CAPEX has been kept unchanged, the efficiency and the battery capacity, \ie the range, have been increased. As seen in the results, this change of data made \gls{BEV} often more competitive than its hydrogen-based equivalent.

\section{End-use demands}
\label{sec:cs:demand}
End-use demands, exogenously imposed as inputs to the model, are characterised by yearly quantities to satisfy and are also distributed over the different hours of each representative year of the transition, to account for their daily or seasonal variability \cite{Limpens2020,limpens2021generating}. In this work, the yearly end-use demands (EUD) for all sectors are calculated from the forecast proposed by the European Commission for Belgium (Appendix 2 in report \cite{EuropeanCommission2021}).

\subsection{Non-energy demand}
\label{subsec:cs:NED}
The \gls{NED} currently represents around 20\% of the final energy consumption in Belgium \cite{FPSEconomy2021}.  This section summarises the rationale of adding a higher level of details to the \gls{NED} compared to what was done in the previous version of the case study \cite{limpens2021generating}. Then, it explains the methodology used to quantify this demand.\\

@@

Over the recent history, there has been a relatively constant share of three main categories of the final consumption for non-energy use in Belgium \cite{statbel_NED_2019} (see Figure \ref{fig:cs_NED_hist}): (i) naphtha and \gls{LPG} (between 59\% and 67\% of the total final consumption, around 59.4\,TWh in 2019), (ii) fossil gas (between 9\% and 14\%, 11.8\,TWh in 2019), and (iii) others (\ie bitumen, coal tar and other oil products) (between 21\% and 28\%).

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.8\textwidth]{NED_hist.pdf}
\caption{Historical data for the non-energy use in Belgium \cite{statbel_NED_2019}. ``Other oil products" account for tar and sulphur. This group also includes by-products of the oil-refining industry like aromatics (\eg \acrfull{BTX}) and olefins (\eg propylene). Graph adapted from \cite{rixhon2021comprehensive}.}
\label{fig:cs_NED_hist}
\end{figure}

Naphtha and \gls{LPG} are consumed in a naphtha cracker, which results in ethylene and propylene, what will be considered as \gls{HVC} in the rest of this work. Similarly, fossil gas, as non-energy carrier, is used in steam methane reformers to produce the required hydrogen for the synthesis of ammonia. The small shares of bitumen and coal tar are used for roadworks and to produce synthetic gas through gasification, respectively. Finally \og other oil products'' take into account, indistinguishably, tar and sulphur as well as by-products of the refineries (\eg \gls{BTX}). About methanol, there is currently no production plant in Belgium even if the country plays a role in trading this commodity between its neighbouring countries and consumes part of what it imports.\\

@@

\noindent
The non-energy demand studied in this work focuses on the chemical industry (more than 90\% of the non-energy use in Belgium) and, similarly to other studies \cite{IEA2018_petrochemicals, daioglou2014energy}, is split between the three aforementioned main groups of products (\ie \gls{HVC}, ammonia and methanol). Before describing these three demands, this study excludes bitumen, coal tar and \og other oil products\fg. The first two represent marginal shares of the current non-energy use (\ie 4\% and 5\% respectively) in such a way that they should not affect the conclusions. As described previously, the latest are mostly by-products from refineries that the system uses because they are available. However, in a perspective of defossilisation, since the future of fossil-based refineries is unclear, they have not been implemented in this study nor their by-products.

Regarding \gls{HVC}, the future of their production is highly uncertain. Because of the new regulations and strategies promoting recycling and limitation of single-use plastics \cite{EU_plastics}.  Besides this uncertainty, Belgium stays a major exporter as approximately 2/3 of plastic raw materials produced locally are exported abroad \cite{agoria_plastics}. Even if a significant part of \gls{HVC} produced in Belgium is not locally consumed, this demand has been set based on the assumption that Belgium will keep its industrial activity in this sector. In other words, we do not deduce the part of local production of \gls{HVC} being then exported (and not consumed locally), unlike ammonia and methanol, which will be more traded commodities in the future (as energy carriers and non-energy products). Therefore, the actual demand of \gls{HVC} is inferred from the consumption of naphtha and \gls{LPG} as non-energy use as well as energy-carrier in the chemical and petrochemical industries \cite{statbel_NED_2019}. This assumption is based on the fact that, in the conversion processes to produce \gls{HVC} from naphtha or \gls{LPG}, these fuels also serve as energy-carrier to supply the processes themselves. Then, given the respective efficiencies (1.83t$_{\text{naphtha}}$/t$_{\text{HVC}}$ and 1.67t$_{\text{LPG}}$/t$_{\text{HVC}}$) \cite{IEA2018_petrochemicals}, the current yearly demand of \gls{HVC} is estimated equal to 3069\,kt, without making distinctions between the different chemicals (\ie ethylene, propylene and \gls{BTX}).

The ammonia sector in Belgium is quite different: the country locally produces and imports ammonia much more than it exports it. Thanks to a database from the United Nations \cite{UN_statistics} and the National Bank of Belgium, it has been identified that, over the last ten years, Belgium, on average, has imported 1010\,kt of ammonia, exported 105\,kt and locally produced 990\,kt per year. Therefore, on top of the local production, the net import (\ie import minus export) is also included in this non-energy demand. This gives a current yearly demand of 1895\,kt of ammonia.

Concerning the demand of methanol, similarly to ammonia, this work solely considers the net imports as there is no local production in Belgium. To define the actual non-energy demand of methanol, only a 51\%-share of this net import is kept since, according to the Methanol Institute and \gls{MMSA}, this share is used for formaldehyde production in Belgium \cite{MMSA51}. Currently, the rest of the methanol is used for energy purposes, mostly as \gls{MTBE} in gasoline blending. This methodology gives a current yearly non-energy demand of methanol of 269\,kt.

Eventually, after converting these masses of products into energy contents (\ie LHV: \gls{HVC} - 47\,MJ/kg, ammonia - 18.8\,MJ/kg and methanol - 19.9\,MJ/kg), we assume that the current shares of each of the three commodities (\ie 77.9\%, 19.2\% and 2.9\% for \gls{HVC}, ammonia and methanol, respectively) are supposed to remain unchanged over the transition.

\subsection{Forecast of the demands over the transition}
\label{subsec:cs:EUD_forecast}
In its latest report \cite{EuropeanCommission2021}, the European Commission forecasts a significant and abrupt increase of the \gls{NED} compared to their previous report \cite{EuropeanCommission2016}, \ie +80\% over the 2020-2030 time window. Given this discrepancy that is unsubstantiated and specific to the case of Belgium, the evolution trend of the \gls{NED} of the current work has been inferred from the previous edition, published in 2016, \cite{EuropeanCommission2016}. Between 2020 and 2050, one observes a noteworthy increase of the electricity (+40\%), passenger (+45\%) and freight mobility (+35\%) demands (see \Cref{fig:cs_demands}). The rise of the non-energy demand is more limited, \ie +6\%, whereas the heating demands is forecast to decrease: -11\% for the low-temperature heat demand and -3\% for the high-temperature heat demand. This is explained by a better insulation of buildings and an improved efficiency of industrial processes. Regarding the top-right graph of \Cref{fig:cs_demands}, it is the aggregation of the same data as in the top-left graph but per economic sector, rather than per energy sector, with the non-energy demand being associated with the industry. This illustrates how industrialised Belgium is, compared to households and services, and, consequently, highly energy-intensive. The bottom graph of \Cref{fig:cs_demands} gives the passenger and the freight mobility. The sharp increase from 2020 to 2025 is due the COVID-crisis that led to significantly reduced demands in 2020. As far as the hourly discretisation of these demands is concerned, time series are based on historical values of 2015 for the fluctuating parts of electricity and low-temperature heating demands \cite{Limpens2020}. A daily time series is used for the passenger mobility and applied similarly to every days of the year. Finally, for the other demands (\ie high-temperature heat, freight mobility, \gls{NED} and the share of electricity and low-temperature heat demands that are considered as constant over the year), the yearly demand is distributed uniformly over the different hours of the year.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.49\textwidth]{EUD_sec.pdf}
\includegraphics[width=0.49\textwidth]{EUD_cat.pdf}
\includegraphics[width=0.49\textwidth]{EUD_mob.pdf}
\caption{EnergyScope splits the whole-energy system end-use demands (EUD) into two sets: (non)-energy and transport-related. This figure presents the yearly values of each of these demands. In the top-right graph, the non-energy demand has been fully associated with the industrial demand. This highlights the significant level of industrialisation of Belgium compared to the other sectors.}
\label{fig:cs_demands}
\end{figure}

\section{Resources}
\label{sec:cs:resources}
To supply the aforementioned demands, EnergyScope Pathway implements a variety of resources defined by their cost of purchasing, $\mathit{c}_{\mathrm{op}}$, their global warming potential, $\emph{gwp}_{\mathrm{op}}$, as well as their 
availability, as detailed by \citet{limpens2024pathway}. First, the evolution of the respective costs of purchasing is presented (see \Cref{fig:cs_resources_cost}). Regarding ``renewable electrofuels'', the costs are in line with the recent study of \citet{genge2023supply} who carried out an extensive review and ``meta-analysis \cite{grant2009typology,page2021prisma} of 30 studies on the supply costs of chemical energy carriers''. Then, besides their costs, the resources are either limited or unlimited in terms of availability and either renewable or not. The limitation in terms of availability can be direct or indirect. On the one hand, woody (23.4\,TWh) and wet biomass (38.9\,TWh) are directly limited by their local potentials and the consumption of waste (17.8\,TWh) and coal (33.4\,TWh) is assumed not to exceed the current use. The local potential of biomass, \ie about 60\,TWh in total, could be considered as optimistic given the work of \citet{colla2024navigating} rather limiting this potential to about 40\,TWh. On the other hand, wind, solar, hydro and uranium are limited by the technical potentials of \gls{PV} panels (59.2\,GW), onshore (10\,GW) and offshore (6\,GW) wind turbines, run-of-the-river power plants (0.1\,GW) and the choice to limit nuclear power plants to 6\,GW, respectively. In line with the work of \citet{PATHS2050} and the maximum capacity of conventional nuclear reactors that have been installed in Belgium, the same 6\,GW are assumed to be the maximum capacity for \gls{SMR}. Imported electricity is limited in two ways: the potential of instantaneous capacity of interconnection with neighbouring countries (\ie 11.9\,GW by 2050 \cite{ELIA_2050}) and a limitation to 30\% of the yearly electricity end-use demand (\ie 32.4\,TWh by 2050) \cite{limpens2021generating}. In the current work, the electrofuels (\ie e-methane, e-hydrogen, e-methanol and e-ammonia) are assumed to be ``sustainable" in the sense that they do not increase the concentration of \ce{CO2} in the atmosphere \cite{rixhon2021terminology}. In practice, it means that their \gls{GWP} is assumed to be zero in the model. Regarding these electrofuels, the \citet{h2coalition} has carried out an extensive techno-economic analysis to estimate their respective cost of purchasing, after having identified some key locations from which importing these energy carriers (\eg Chile, Australia or Morocco). As the amount to import from each of these locations is hard to forecast, the current work considers the average cost between the different locations. Besides these, every other resource has its specific \gls{GWP} like coal ($\emph{gwp}_{\mathrm{op,coal}}=0.40$\,kt$_{\ce{CO2},\text{eq}}$/GWh), natural gas ($\emph{gwp}_{\mathrm{op,NG}}=0.27$\,kt$_{\ce{CO2},\text{eq}}$/GWh) or the fossil-based molecules equivalent to the electrofuels (\eg $\emph{gwp}_{\mathrm{op,ammonia}}=0.46$\,kt$_{\ce{CO2},\text{eq}}$/GWh or $\emph{gwp}_{\mathrm{op,methanol}}=0.41$\,kt$_{\ce{CO2},\text{eq}}$/GWh).

@@

\includegraphics[width=0.49\textwidth]{Res_ren.pdf}
\includegraphics[width=0.49\textwidth]{Res_foss.pdf}
\includegraphics[width=0.49\textwidth]{Res_others.pdf}
\caption{Cost of purchasing the different resources. Besides the free local renewables (\ie sun, wind and hydro) limited by technical potentials, EnergyScope accounts for renewable energy carriers and their respective fossil counterparts (top-left and top-right graphs). These fuels can be imported from abroad without limitation on their availability. Other carriers are limited either by their local potentials (\ie biomass and waste) or other considerations like the power grid interconnections or the capacity of nuclear power plants.}
\label{fig:cs_resources_cost}
\end{figure}

\section{Conversion technologies}
\label{sec:cs:technologies}
As the end-use demands are defined as energy (and non-energy with the \gls{NED}) services rather than a certain quantity of oil or solar irradiance, for instance, technologies are implemented to convert these resources into the end-use demands. Besides their CAPEX, OPEX and lifetime defined in Section \ref{sec:meth:ES}, production and conversion technologies (\ie \gls{CCGT}, car or boiler) have a conversion efficiency whereas storage technologies (\ie thermal storage, battery or molecule storage) have charge/discharge losses. There are also infrastructure technologies. They encompass, for instance, the power grid, the \gls{DHN} or technologies to produce intermediate energy carriers (\eg wood pyrolysis, biomethanolation or steam methane reforming to produce hydrogen). Not digging into too much details about the exhaustive list of these technologies presented in a previous work \cite{limpens2021generating}, this section rather focuses on the implementation of \glsxtrfull{SMR} and the technologies to supply the \glsxtrfull{NED}.

@@

For further details on these analyses, the interested reader is invited to refer to previously published works \cite{rixhon2021comprehensive,rixhon2022integration}.

@@

The exhaustive list of the parameters accounted for in this work is presented in Appendix \ref{app:UC_full}.
